---
title: Team contract
---

# Rationale

Teams work better when members have a common understanding of the team's goals and expectations for collaboration.
The purpose of this activity is to help your team making a plan for working together during lab and outside of the scheduled lab time.
Each team member will have some ideas about how a team should operate.
These ideas may be very different.
This is your opportunity to share your thoughts and ideas to promote optimal team function and prevent misunderstandings in the future.

# Instructions

Discuss each of the items below as a team.
Have one person act as the **recorder** and type the team's decisions in this document.
*Be sure the team agrees on an item before it is added to the document.*

Once the document is complete, the recorder should render, commit, and push the team agreement to GitHub.
All team members can refer to this document throughout the semester.

# Goals

-   What are your team goals for this project?

To develop an astute understanding of the statistical models that shape the world around us.

-   What do you want to accomplish? What skills do you want to develop or refine?

Collaboration, problem-solving, communication, staying calm, resolving issues with grace, coordination, writing good commit messages, being clear and organized

# Expectations

Don't be passive aggressive but offer constructive criticism. Comunicate. If you don't understand something, reach out. It's okay to make mistakes, just communicate.

## Meetings

Tuesdays at 3 work for all of us.

Facetime/zoom will be our meeting location.

## Communication

We made a IMessage groupchat We'll text in the chat if we can't make it to lab.


## Workflow

We decided to finish our work 2 days before the actual deadline set by the class to figure out conflicts and clean up code and writing.

Text in the chat, ask for help, don't be shy! Everyone has workloads spike up at different times.


# Policies & procedures

Communicate, talk to each other, don't struggle in silence. If something's bothering you, speak up before it turns into a bigger issue.

# Consequences

We will communicate honestly and openly but not passively-aggressively. If things get bad, we'll talk to a TA to help us sort out.

# Additional items

We want to plan out a dataset that everyone finds fun and everyone will enjoy to work on. We want this project to be fun rather than a burden. We want to work effectively as a team!!!
--------------------------------------------------------------------------------

We share these goals and expectations, and agree to these policies, procedures, and consequences.

Team member name: Camden Chin 

Team member name: Lexi Payne

Team member name: David Boham

Team member name: Isha Omer

...



```{r}
library(tidyverse)
library(janitor)
library(stringr)
library(lubridate)

#import movies data
movies_cleaned <- read_csv("data/Movies - Cleaned.csv") |> clean_names()
movies_sequels <- read_csv("data/Movies with sequels.csv") |> clean_names()

```



```{r}
#| label: join-data-movies_cleaned_and_movies_sequels
#joined both datasets from "data 1" from project proposal

joined_sequel_data <- movies_cleaned |>
  left_join(
    movies_sequels,
    by = c("title", "year"),
    suffix = c("_cleaned", "_sequels")
  )

joined_sequel_data
```

```{r}
#| label: import-imdb-data

imdb_movies <- read_csv("data/IMDb movies.csv") |> clean_names()
imdb_ratings <- read_csv("data/IMDb ratings.csv") |> clean_names()

```
```{r}
#| label: join-imdb-data-info
#joined both datasets from "data 2" from project proposal

imdb_full <- imdb_movies |>
  left_join(imdb_ratings, by = "imdb_title_id")

```

```{r}
#| label: join-both-datasets
#| 
final_joined_data <- joined_sequel_data |>
  left_join(
    imdb_full,
    by = c("title", "year"),
    suffix = c("_sequels", "_imdb")
  )

final_joined_data
```

