---
title: "Sequels: More or Less Successful"
subtitle: Report
author: 
  - name: Isha Omer
  - name: Lexi Payne
  - name: Camden Chin
  - name: David Boham
execute:
  echo: False
---

Add 1-3 sentence project abstract here.

## Introduction

This project investigates whether movie sequels are more or less successful than 
their originals. Using two datasets — a curated list of movies with sequels and 
IMDb ratings and financial data — we explore how sequel performance compares 
across metrics like ratings and box office revenue. All data was sourced from
IMDb and compiled for the STA 199 course project.

## Data 
```{r}
#| label: load-libraries
#| message: false
#| warning: false
library(tidyverse)
library(janitor)
library(stringr)
library(lubridate)
```

```{r}
#| label: import-data
#| message: false
#| warning: false
movies_cleaned <- read_csv("data/Movies - Cleaned.csv") |> clean_names()
movies_sequels <- read_csv("data/Movies with sequels.csv") |> clean_names()
imdb_movies    <- read_csv("data/IMDb movies.csv")        |> clean_names()
imdb_ratings   <- read_csv("data/IMDb ratings.csv")       |> clean_names()
```

## Data Cleaning

After importing the raw datasets, we clean and standardize each one to ensure 
consistency before joining. This includes fixing data types, removing duplicates, 
and dropping rows with missing values in key fields. Note that `clean_names()` 
was already applied on import, so it is not repeated here.
```{r}
#| label: clean-all-datasets
#| message: false
#| warning: false

# Clean movies_cleaned
movies_cleaned <- movies_cleaned |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean movies_sequels
movies_sequels <- movies_sequels |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_movies
imdb_movies <- imdb_movies |>
  distinct() |>
  mutate(
    title                 = str_squish(str_to_title(title)),
    year                  = as.integer(year),
    budget                = as.numeric(str_replace_all(budget, "[^0-9.]", "")),
    usa_gross_income      = as.numeric(str_replace_all(usa_gross_income, "[^0-9.]", "")),
    worlwide_gross_income = as.numeric(str_replace_all(worlwide_gross_income, "[^0-9.]", ""))
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_ratings
imdb_ratings <- imdb_ratings |>
  distinct() |>
  filter(!is.na(imdb_title_id), !is.na(weighted_average_vote))
```

## Data Joining

Here we join the two movie datasets (Data Source 1 from the project proposal) 
using a left join on both `title` and `year` to ensure we're matching the 
correct films. The suffix argument distinguishes columns that appear in both 
datasets. The result, `joined_sequel_data`, retains all rows from 
`movies_cleaned` and appends sequel information where available.
```{r}
#| label: join-sequel-data
#| message: false

joined_sequel_data <- movies_cleaned |>
  left_join(
    movies_sequels,
    by     = c("title", "year"),
    suffix = c("_cleaned", "_sequels")
  )

head(joined_sequel_data, 5)
```

The two IMDb datasets are joined using `imdb_title_id` as the key, which is a 
unique identifier shared across both files. This produces a single IMDb dataset 
`imdb_full` that contains both movie metadata and ratings information.
```{r}
#| label: join-imdb-data
#| message: false
imdb_full <- imdb_movies |>
  left_join(imdb_ratings, by = "imdb_title_id")
```

Finally, we combine both data sources into one unified dataset. Because some 
titles exist in IMDb under a different release year, we join on title only and 
then keep the closest year match within a 2-year window. This significantly 
improves the number of successful matches over a strict title + year join.
```{r}
#| label: final-join
#| message: false
#| warning: false
final_joined_data <- joined_sequel_data |>
  left_join(
    imdb_full,
    by           = "title",
    suffix       = c("_sequels", "_imdb"),
    relationship = "many-to-many"
  ) |>
  mutate(year_diff = abs(year_sequels - year_imdb)) |>
  group_by(title, year_sequels) |>
  slice_min(year_diff, n = 1) |>
  ungroup() |>
  filter(year_diff <= 2)

head(final_joined_data, 5)
```

## Join Summary

All 272 movies successfully matched with IMDb data, confirming that the 
title-based join with a 2-year tolerance worked correctly. This gives us a 
complete dataset to work with for the rest of the analysis.
```{r}
#| label: join-summary
#| message: false
final_joined_data |>
  summarise(
    matched   = sum(!is.na(imdb_title_id)),
    unmatched = sum(is.na(imdb_title_id))
  )
```

## Creating New Columns

Since sequels often have much larger production budgets, simply comparing total income might not be a fair comparison. A movie with a huge budget might naturally bring in more income because of marketing, special effects, or brand recognition. Therefore, dividing income/budget to create a income-to-budget ratio will allow comparisons on how successful films are relative to how much money was invested in them.

```{r}
#| label: creating-revenue-budget-ratio

final_joined_data <- final_joined_data |> 
  mutate(income_budget_ratio = worlwide_gross_income / budget)
```


## Analysis

To analyze whether sequels or original films performed better, we investigated two main variables: financial success and IMDb ratings.

Financial success was the first variable measured. Again, it was computed by taking the ratio of income to budget, so financial success would be measured relative to the amount of money put into the film.

```{r}
#| label: financial-success-plot
#| warning: FALSE

final_joined_data <- final_joined_data |>
  filter(!is.na(order)) |>  # keep only movies that are part of a series
  mutate(
    sequel_status = if_else(order == 1, "Original", "Sequel_or_later")
  )

final_joined_data |>
  ggplot(aes(x = income_budget_ratio, y = sequel_status , fill = sequel_status)) +
  geom_boxplot(alpha = 0.7) +
  scale_x_log10() +
  labs(
    x = "Worldwide income / budget (log scale)",
    y = "Movie type",
    title = "Relative Financial Success (income/budget): Originals vs Sequels"
  ) +
  theme_minimal() +
  scale_y_discrete(
    labels = c("Original", "Sequel or Later")
  ) +
  theme(legend.position = "none")
```


```{r}
#| label: budget-median

final_joined_data |> 
  group_by(sequel_status) |> 
  summarize(
    median = median(income_budget_ratio, na.rm = TRUE)
  )
```

The median income-to-budget ratio for original movies was about 4, while that of sequels or later films was about 3.7. Thus, a slight increase in the median financial success can be seen in original films compared to moves that succeed them. This same result is also seen in the graph, where the median of the income/budget ratio for original films is slightly higher than that of later films. However, it is important to note that the boxplots for both original and sequel-or-later films have outliers on both the higher and lower ends, showing there are movies in both categories that underperform and overperform. The distribution of original films' outliers shows more variability (in other words, the outliers span a greater distance), while the distribution of sequel-or-later films' outliers span a smaller distance, suggesting more predictable outcomes in seqeul-or-later films.

Overall, the graph and summary statistics suggest original films have slightly greater financial success than sequel-or-later films.


IMDb rating was the next variable analyzed.

```{r}
#| label: rating-plot

final_joined_data |>
  ggplot(aes(x = im_db_rating_cleaned, y = sequel_status , fill = sequel_status)) +
  geom_boxplot(alpha = 0.7) +
  scale_x_log10() +
  labs(
    x = "IMDb Rating",
    y = "Movie type",
    title = "Relative Rating Success: Originals vs Sequels"
  ) +
  theme_minimal() +
  scale_y_discrete(
    labels = c("Original", "Sequel or Later")
  ) +
  theme(legend.position = "none")
```

```{r}
#| label: rating-median

final_joined_data |> 
  group_by(sequel_status) |> 
  summarize(
    median = median(im_db_rating_cleaned)
  )
```

The median IMBDb rating of original films (6.9 out of 10) was higher for original films than sequel-or-later films (6.2 out of 10). This same result is also seen in the graph, where the median of the IMDb ratings for original films is slightly higher than that of later films. It's also important to note that the distribution of ratings for sequel-or-later films seems to have greater variability; aditionally, while both distributions have outliers, the sequel-or-later films have 4 more outliers than the original films distribution, and they are also lower, suggesting that sequel-or-later films tend to have a harder time maintaining high audience enjoyment compared to original films.

Overall, the graph and summary statistics suggest original films have slightly greater IMDd ratings than sequel-or-later films.



```{r}
#| label: plot-of-budget-vs-income
#| fig-width: 8
#| fig-asp: 0.7
#| warning: false


ggplot(final_joined_data,
       aes(x = budget, y = worlwide_gross_income, color = sequel_status)) +
  geom_point(alpha = 0.7) +
  labs(
    title = "Budget vs Revenue",
    subtitle = "Of Original Movies vs Sequels",
    x = "Budget",
    y = "Worldwide Gross Income",
    color = "Sequel Status"
  ) +
  geom_smooth(se = FALSE, method = "lm", linewidth = 0.5)
```


## Discussion

Add discussion section here.
