---
title: "Sequels: More or Less Successful"
subtitle: Report
author: 
  - name: Isha Omer
  - name: Lexi Payne
  - name: Camden Chin
  - name: David Boham
execute:
  echo: False
  warning: False
---

# Project Abstract
Our project investigates the relative success of movie sequels compared to the original films by analyzing financial performance and IMDb ratings across 272 movies. We standardized financial success by creating an income-to-budget ratio metric, where we found that original films (median ratio ~ 4.0) slightly outperform sequels (median ratio ~3.7) and maintain higher audience ratings (6.9 vs. 6.2). Our findings suggest that despite the expectation of the pre-existing audience of franchises, original productions tend to achieve slightly greater  profitability.

## Introduction

This project investigates whether movie sequels are more or less successful than 
their originals. Using two data sets — a curated list of movies with sequels and 
IMDb ratings and financial data — our research question is, do sequels or originals perform better
across metrics like ratings and box office revenue when accounting for budget?

Project motivations were both silly and serious. All of us are big movie lovers and wanted to do something fun and lighthearted and lean into our hobbies. Additionally, one of us hopes to pursue film making after graduation. More "serious" motivations for the project include wanting to see if there are statistical bases for phenomena such as "franchise fatigue" and wanting to explore the risks filmmakers have to consider when releasing originals vs. sequels.

All data was sourced from IMDb and compiled for the STA 199 course project. Where users submit ratings and professionals/scrapers provide financial figures. The data sets were taken from Kaggle:

[IMDb movie ratings data set](https://www.kaggle.com/datasets/sharathmudigoudr/imdb-movie-dataset-from-year-1893-to-2020/data)

[Sequel data set set](https://www.kaggle.com/code/cpbdev/sequels-dataset/input)

## Data 
```{r}
#| label: load-libraries
#| message: false
#| warning: false
library(tidyverse)
library(janitor)
library(stringr)
library(lubridate)
```

```{r}
#| label: import-data
#| message: false
#| warning: false
movies_cleaned <- read_csv("data/Movies - Cleaned.csv") |> clean_names()
movies_sequels <- read_csv("data/Movies with sequels.csv") |> clean_names()
imdb_movies    <- read_csv("data/IMDb movies.csv")        |> clean_names()
imdb_ratings   <- read_csv("data/IMDb ratings.csv")       |> clean_names()
```

## Data Cleaning

After importing the raw datasets, we clean and standardize each one to ensure 
consistency before joining. This includes fixing data types, removing duplicates, 
and dropping rows with missing values in key fields. Note that `clean_names()` 
was already applied on import, so it is not repeated here.
```{r}
#| label: clean-all-datasets
#| message: false
#| warning: false

# Clean movies_cleaned
movies_cleaned <- movies_cleaned |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean movies_sequels
movies_sequels <- movies_sequels |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_movies
imdb_movies <- imdb_movies |>
  distinct() |>
  mutate(
    title                 = str_squish(str_to_title(title)),
    year                  = as.integer(year),
    budget                = as.numeric(str_replace_all(budget, "[^0-9.]", "")),
    usa_gross_income      = as.numeric(str_replace_all(usa_gross_income, "[^0-9.]", "")),
    worlwide_gross_income = as.numeric(str_replace_all(worlwide_gross_income, "[^0-9.]", ""))
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_ratings
imdb_ratings <- imdb_ratings |>
  distinct() |>
  filter(!is.na(imdb_title_id), !is.na(weighted_average_vote))
```

## Data Joining

Here we join the two movie datasets (Data Source 1 from the project proposal) 
using a left join on both `title` and `year` to ensure we're matching the 
correct films. The suffix argument distinguishes columns that appear in both 
datasets. The result, `joined_sequel_data`, retains all rows from 
`movies_cleaned` and appends sequel information where available.
```{r}
#| label: join-sequel-data
#| message: false

joined_sequel_data <- movies_cleaned |>
  left_join(
    movies_sequels,
    by     = c("title", "year"),
    suffix = c("_cleaned", "_sequels")
  )

```

The two IMDb datasets are joined using `imdb_title_id` as the key, which is a 
unique identifier shared across both files. This produces a single IMDb dataset 
`imdb_full` that contains both movie metadata and ratings information.
```{r}
#| label: join-imdb-data
#| message: false
imdb_full <- imdb_movies |>
  left_join(imdb_ratings, by = "imdb_title_id")
```

Finally, we combine both data sources into one unified dataset. Because some 
titles exist in IMDb under a different release year, we join on title only and 
then keep the closest year match within a 2-year window. This significantly 
improves the number of successful matches over a strict title + year join.
```{r}
#| label: final-join
#| message: false
#| warning: false
final_joined_data <- joined_sequel_data |>
  left_join(
    imdb_full,
    by           = "title",
    suffix       = c("_sequels", "_imdb"),
    relationship = "many-to-many"
  ) |>
  mutate(year_diff = abs(year_sequels - year_imdb)) |>
  group_by(title, year_sequels) |>
  slice_min(year_diff, n = 1) |>
  ungroup() |>
  filter(year_diff <= 2)

```
## Data Definitions

* **`title`**: Official name of the movie, used as the primary key for joining datasets.
* **`year`**: Release year, used to guarantee correct matching between sequels and originals.
* **`budget`**: Production cost of the movie.
* **`worlwide_gross_income`**: Total revenue generated globally.
* **`income_budget_ratio`**: A self-derived metric calculated as $\frac{\text{worldwide\_gross\_income}}{\text{budget}}$. This standardizes financial success relative to the initial budget.
* **`im_db_rating_cleaned`**: Weighted average of user vote from IMDb on a scale of 1–10.
* **`sequel_status`**: Labels films as either "Original" or "Sequel or Later".

## Join Summary

All 272 movies successfully matched with IMDb data, confirming that the 
title-based join with a 2-year tolerance worked correctly. This gives us a 
complete dataset to work with for the rest of the analysis.
```{r}
#| label: join-summary
#| message: false

invisible(
  final_joined_data |>
    summarise(
      matched   = sum(!is.na(imdb_title_id)),
      unmatched = sum(is.na(imdb_title_id))
  )
)
```

## Creating New Columns

Since sequels often have much larger production budgets, simply comparing total income might not be a fair comparison. A movie with a huge budget might naturally bring in more income because of marketing, special effects, or brand recognition. Therefore, dividing income/budget to create a income-to-budget ratio will allow comparisons on how successful films are relative to how much money was invested in them.

```{r}
#| label: creating-revenue-budget-ratio

final_joined_data <- final_joined_data |> 
  mutate(income_budget_ratio = worlwide_gross_income / budget)


```


## Analysis

To analyze whether sequels or original films performed better, we investigated two main variables: financial success and IMDb ratings.

Financial success was the first variable measured. Again, it was computed by taking the ratio of income to budget, so financial success would be measured relative to the amount of money put into the film.

```{r}
#| label: financial-success-plot
#| warning: FALSE

final_joined_data <- final_joined_data |>
  filter(!is.na(order)) |>  # keep only movies that are part of a series
  mutate(
    sequel_status = if_else(order == 1, "Original", "Sequel_or_later")
  )

final_joined_data |>
  ggplot(aes(x = income_budget_ratio, y = sequel_status , fill = sequel_status)) +
  geom_boxplot(alpha = 0.7) +
  scale_x_log10() +
  labs(
    x = "Worldwide income / budget (log scale)",
    y = "Movie type",
    title = "Relative Financial Success (income/budget):
    Originals vs Sequels"
  ) +
  theme_minimal() +
  scale_y_discrete(
    labels = c("Original", "Sequel or Later")
  ) +
  theme(legend.position = "none")
```


```{r}
#| label: budget-median

final_joined_data |> 
  group_by(sequel_status) |> 
  summarize(
    Median = median(income_budget_ratio, na.rm = TRUE)
  ) |> 
  mutate(
    sequel_status = stringr::str_replace_all(sequel_status, "_", " "),
    sequel_status = stringr::str_to_title(sequel_status)
  ) |> 
   rename(
    "Sequel Status" = sequel_status,
    "Median" = Median
  ) |> 
  knitr::kable(digits = 2)
```

The median income-to-budget ratio for original movies was about 4, while that of sequels or later films was about 3.7. Thus, a slight increase in the median financial success can be seen in original films compared to moves that succeed them. This same result is also seen in the graph, where the median of the income/budget ratio for original films is slightly higher than that of later films. However, it is important to note that the boxplots for both original and sequel-or-later films have outliers on both the higher and lower ends, showing there are movies in both categories that underperform and overperform. The distribution of original films' outliers shows more variability (in other words, the outliers span a greater distance), while the distribution of sequel-or-later films' outliers span a smaller distance, suggesting more predictable outcomes in seqeul-or-later films.

Overall, the graph and summary statistics suggest original films have slightly greater financial success than sequel-or-later films.


IMDb rating was the next variable analyzed.

```{r}
#| label: rating-plot

final_joined_data |>
  ggplot(aes(x = im_db_rating_cleaned, y = sequel_status , fill = sequel_status)) +
  geom_boxplot(alpha = 0.7) +
  scale_x_log10() +
  labs(
    x = "IMDb Rating",
    y = "Movie type",
    title = "Relative Rating Success: Originals vs Sequels"
  ) +
  theme_minimal() +
  scale_y_discrete(
    labels = c("Original", "Sequel or Later")
  ) +
  theme(legend.position = "none")
```

```{r}
#| label: rating-median

final_joined_data |> 
  group_by(sequel_status) |> 
  summarize(
    Median = median(im_db_rating_cleaned)
  ) |> 
  mutate(
    sequel_status = stringr::str_replace_all(sequel_status, "_", " "),
    sequel_status = stringr::str_to_title(sequel_status)
  ) |> 
  rename(
    "Sequel Status" = sequel_status,
    "Median" = Median
  ) |> 
  knitr::kable(digits = 3)
```

The median IMBDb rating of original films (6.9 out of 10) was higher for original films than sequel-or-later films (6.2 out of 10). This same result is also seen in the graph, where the median of the IMDb ratings for original films is slightly higher than that of later films. It's also important to note that the distribution of ratings for sequel-or-later films seems to have greater variability, which suggests  less consistent audience reception; additionally, while both distributions have outliers, the sequel-or-later films have 4 more outliers than the original films distribution, and they are also lower, suggesting that sequel-or-later films tend to have a harder time maintaining high audience enjoyment compared to original films. Nevertheless, both boxplots do overlap indicating that some sequels can still be successful.

Overall, the graph and summary statistics suggest original films have slightly greater IMDd ratings than sequel-or-later films.


## Discussion

  This study investigated whether movie sequels/succeeding films or original films seem to have greater success. Success was broken down into two categories - financial success and rating success. It was found that original films have greater success in both categories.
  
  Financial success was calculated as an income-to-budget ratio, and the median for original films, around 4, was slightly higher than that of sequel-or-later films, about 3.7. In other words, original films, on average, earn more money relative to money invested. This suggests that even though sequel-or-later films often have a "fan base," or a group of people invested in the first film and wanting to watch more, this doesn't often lead to greater financial success. Thus, even though original films are "riskier" in the sense that filmmakers have no basis to conclude how their film might be perceived by audiences, they still tend to result in more income. This could influence how directors and creators of films choose which projects to take on.
  
  IMDb ratings were used to measure rating success, and again, the median IMDb rating for original films (6.9 out of 10) was higher than that of sequel-or-later films (6.2 out of 10). Factors that can explain decreased ratings for sequel-or-later films include a decline in the novelty of the film's plot, higher expectations that the first film established leading to a sense of being let down, and the phenomenon of "franchise fatigue," where viewers no longer look forward to a continuation of a story due to issues such as decreasing quality or lack of originality.
  
  The analysis includes several limitations. One is that all films that came after the original were put into one category. In other words, if a movie franchise has 3 movies in the series, the first one, the original one, is given its own category, while the sequel and third movie are put into the same category. Thus, nuances may be lost (for example, a sequel may have performed really well, but the third installment may have performed very poorly, making the overall "sequels-or-later" category for that franchise have a low financial/rating success). A possible improvement to address this issue is to show these nuances by comparing the financial and rating success for each film in a sequence to its original separately.
  
  Another limitation is that factors like genre may have influenced whether sequels seem to have more success than their original films, but these factors were not taken into account in the analysis. A possible improvement to address this issue is to repeat the analysis but break up the graphs and summary statistics into categories, like genre.
  
  A final limitation is the use of IMDb ratings as an indicator for how the movies were perceived by the audience. IMDb ratings may be susceptible to problems like response bias and may not be representative of the actual quality of the movie if users are voting based on other factors, such as rating movies higher if it includes their favorite actors. A possible improvement to address this issue is that other variables could be taken into account to get a more reliable overall rating score, such as by also considering critics' scores.
  
Citations:
[Franchise Fatigue](https://drive.google.com/file/d/1PWhOnjvMQa00je5yCjWHsFqEL2UcRn4q/view)
[IMDb Rating Limitations](https://www.theringer.com/2019/06/12/tv/imdb-rating-system-problems-chernobyl)

  
