---
title: "Sequels: More or Less Successful"
subtitle: Report
author: 
  - name: Isha Omer
  - name: Lexi Payne
  - name: Camden Chin
  - name: David Boham
execute:
  echo: False
  warning: False
---

# Project Abstract
Our project investigates the relative success of movie sequels compared to the original films by analyzing financial performance and IMDb ratings across 272 movies. We standardized financial success by creating an income-to-budget ratio metric, where we found that original films (median ratio ~4.0) slightly outperform sequels (median ratio ~3.7) and maintain higher audience ratings (6.9 vs. 6.2). An installment-level analysis further reveals that while audience ratings decline monotonically across a franchise (6.9 → 6.4 → 6.1), financial returns recover by the third installment, suggesting franchise earning power decouples from audience reception over time. Our findings suggest that despite the expectation of the pre-existing audience of franchises, original productions tend to achieve slightly greater profitability — but that established franchises retain box-office pull largely independent of per-film quality.

## Introduction

This project investigates whether movie sequels are more or less successful than 
their originals. Using two data sets — a curated list of movies with sequels and 
IMDb ratings and financial data — our research question is, do sequels or originals perform better
across metrics like ratings and box office revenue when accounting for budget?

Project motivations were both silly and serious. All of us are big movie lovers and wanted to do something fun and lighthearted and lean into our hobbies. Additionally, one of us hopes to pursue film making after graduation. More "serious" motivations for the project include wanting to see if there are statistical bases for phenomena such as "franchise fatigue" and wanting to explore the risks filmmakers have to consider when releasing originals vs. sequels.

All data was sourced from IMDb and compiled for the STA 199 course project. Where users submit ratings and professionals/scrapers provide financial figures. The data sets were taken from Kaggle:

[IMDb movie ratings data set](https://www.kaggle.com/datasets/sharathmudigoudr/imdb-movie-dataset-from-year-1893-to-2020/data)

[Sequel data set](https://www.kaggle.com/code/cpbdev/sequels-dataset/input)

## Data 

Data sets were taken from Kaggle and were collected by web scraping the IMDb website and putting together information from the IMDb database.
```{r}
#| label: load-libraries
#| message: false
#| warning: false
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(janitor)
library(lubridate)
```

```{r}
#| label: import-data
#| message: false
#| warning: false
movies_cleaned <- read_csv("data/Movies - Cleaned.csv") |> clean_names()
movies_sequels <- read_csv("data/Movies with sequels.csv") |> clean_names()
imdb_movies    <- read_csv("data/IMDb movies.csv")        |> clean_names()
imdb_ratings   <- read_csv("data/IMDb ratings.csv")       |> clean_names()
```

## Data Cleaning

After importing the raw datasets, we clean and standardize each one to ensure 
consistency before joining. This includes fixing data types, removing duplicates, 
and dropping rows with missing values in key fields. Note that `clean_names()` 
was already applied on import, so it is not repeated here.
```{r}
#| label: clean-all-datasets
#| message: false
#| warning: false

# Clean movies_cleaned
movies_cleaned <- movies_cleaned |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean movies_sequels
movies_sequels <- movies_sequels |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_movies
imdb_movies <- imdb_movies |>
  distinct() |>
  mutate(
    title                 = str_squish(str_to_title(title)),
    year                  = as.integer(year),
    budget                = as.numeric(str_replace_all(budget, "[^0-9.]", "")),
    usa_gross_income      = as.numeric(str_replace_all(usa_gross_income, "[^0-9.]", "")),
    worlwide_gross_income = as.numeric(str_replace_all(worlwide_gross_income, "[^0-9.]", ""))
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_ratings
imdb_ratings <- imdb_ratings |>
  distinct() |>
  filter(!is.na(imdb_title_id), !is.na(weighted_average_vote))
```

## Data Joining

Here we join the two movie datasets (Data Source 1 from the project proposal) 
using a left join on both `title` and `year` to ensure we're matching the 
correct films. The suffix argument distinguishes columns that appear in both 
datasets. The result, `joined_sequel_data`, retains all rows from 
`movies_cleaned` and appends sequel information where available.
```{r}
#| label: join-sequel-data
#| message: false

joined_sequel_data <- movies_cleaned |>
  left_join(
    movies_sequels,
    by     = c("title", "year"),
    suffix = c("_cleaned", "_sequels")
  )

```

The two IMDb datasets are joined using `imdb_title_id` as the key, which is a 
unique identifier shared across both files. This produces a single IMDb dataset 
`imdb_full` that contains both movie metadata and ratings information.
```{r}
#| label: join-imdb-data
#| message: false
imdb_full <- imdb_movies |>
  left_join(imdb_ratings, by = "imdb_title_id")
```

Finally, we combine both data sources into one unified dataset. Because some 
titles exist in IMDb under a different release year, we join on title only and 
then keep the closest year match within a 2-year window. This significantly 
improves the number of successful matches over a strict title + year join.
```{r}
#| label: final-join
#| message: false
#| warning: false
final_joined_data <- joined_sequel_data |>
  left_join(
    imdb_full,
    by           = "title",
    suffix       = c("_sequels", "_imdb"),
    relationship = "many-to-many"
  ) |>
  mutate(year_diff = abs(year_sequels - year_imdb)) |>
  group_by(title, year_sequels) |>
  slice_min(year_diff, n = 1) |>
  ungroup() |>
  filter(year_diff <= 2)

```
## Data Definitions

* **`title`**: Official name of the movie, used as the primary key for joining datasets.
* **`year`**: Release year, used to guarantee correct matching between sequels and originals.
* **`budget`**: Production cost of the movie.
* **`worlwide_gross_income`**: Total revenue generated globally.
* **`income_budget_ratio`**: A self-derived metric calculated as $\frac{\text{worldwide\_gross\_income}}{\text{budget}}$. This standardizes financial success relative to the initial budget.
* **`im_db_rating_cleaned`**: Weighted average of user vote from IMDb on a scale of 1–10.
* **`sequel_status`**: Labels films as either "Original" or "Sequel or Later".
* **`installment`**: Groups films by their position in a franchise: 1st (Original), 2nd, or 3rd+.

## Join Summary

All 272 movies successfully matched with IMDb data, confirming that the 
title-based join with a 2-year tolerance worked correctly. This gives us a 
complete dataset to work with for the rest of the analysis.
```{r}
#| label: join-summary
#| message: false

invisible(
  final_joined_data |>
    summarise(
      matched   = sum(!is.na(imdb_title_id)),
      unmatched = sum(is.na(imdb_title_id))
  )
)
```

## Creating New Columns

Since sequels often have much larger production budgets, simply comparing total income might not be a fair comparison. A movie with a huge budget might naturally bring in more income because of marketing, special effects, or brand recognition. Therefore, dividing income/budget to create a income-to-budget ratio will allow comparisons on how successful films are relative to how much money was invested in them.

```{r}
#| label: creating-revenue-budget-ratio

final_joined_data <- final_joined_data |> 
  mutate(income_budget_ratio = worlwide_gross_income / budget)


```


## Analysis

To analyze whether sequels or original films performed better, we investigated two main variables: financial success and IMDb ratings.

Financial success was the first variable measured. Again, it was computed by taking the ratio of income to budget, so financial success would be measured relative to the amount of money put into the film.

```{r}
#| label: financial-success-plot
#| warning: FALSE

all_films <- final_joined_data

final_joined_data <- final_joined_data |>
  filter(!is.na(order)) |>  # keep only movies that are part of a series
  mutate(
    sequel_status = if_else(order == 1, "Original", "Sequel_or_later")
  )

final_joined_data |>
  ggplot(aes(x = income_budget_ratio, y = sequel_status , fill = sequel_status)) +
  geom_boxplot(alpha = 0.7) +
  scale_x_log10() +
  labs(
    x = "Worldwide income / budget (log scale)",
    y = "Movie type",
    title = "Relative Financial Success (income/budget):
    Originals vs Sequels"
  ) +
  theme_minimal() +
  scale_y_discrete(
    labels = c("Original", "Sequel or Later")
  ) +
  theme(legend.position = "none")

```

```{r}
standalone_films <- imdb_full |>
  filter(!is.na(budget), !is.na(worlwide_gross_income),
         budget > 0) |>
  anti_join(final_joined_data, by = "imdb_title_id") |>   # remove franchise films
  mutate(income_budget_ratio = worlwide_gross_income / budget)
```

```{r}
#| label: budget-median

final_joined_data |> 
  group_by(sequel_status) |> 
  summarize(
    Median = median(income_budget_ratio, na.rm = TRUE)
  ) |> 
  mutate(
    sequel_status = stringr::str_replace_all(sequel_status, "_", " "),
    sequel_status = stringr::str_to_title(sequel_status)
  ) |> 
   rename(
    "Sequel Status" = sequel_status,
    "Median" = Median
  ) |> 
  knitr::kable(digits = 2)
```

The median income-to-budget ratio for original movies was about 4, while that of sequels or later films was about 3.7. Thus, a slight increase in the median financial success can be seen in original films compared to moves that succeed them. This same result is also seen in the graph, where the median of the income/budget ratio for original films is slightly higher than that of later films. However, it is important to note that the boxplots for both original and sequel-or-later films have outliers on both the higher and lower ends, showing there are movies in both categories that underperform and overperform. The distribution of original films' outliers shows more variability (in other words, the outliers span a greater distance), while the distribution of sequel-or-later films' outliers span a smaller distance, suggesting more predictable outcomes in seqeul-or-later films.

Overall, the graph and summary statistics suggest original films have slightly greater financial success than sequel-or-later films.


IMDb rating was the next variable analyzed.

```{r}
#| label: rating-plot

final_joined_data |>
  ggplot(aes(x = im_db_rating_cleaned, y = sequel_status , fill = sequel_status)) +
  geom_boxplot(alpha = 0.7) +
  scale_x_log10() +
  labs(
    x = "IMDb Rating",
    y = "Movie type",
    title = "Relative Rating Success: Originals vs Sequels"
  ) +
  theme_minimal() +
  scale_y_discrete(
    labels = c("Original", "Sequel or Later")
  ) +
  theme(legend.position = "none")
```

```{r}
#| label: rating-median

final_joined_data |> 
  group_by(sequel_status) |> 
  summarize(
    Median = median(im_db_rating_cleaned)
  ) |> 
  mutate(
    sequel_status = stringr::str_replace_all(sequel_status, "_", " "),
    sequel_status = stringr::str_to_title(sequel_status)
  ) |> 
  rename(
    "Sequel Status" = sequel_status,
    "Median" = Median
  ) |> 
  knitr::kable(digits = 3)
```

The median IMDb rating of original films (6.9 out of 10) was higher for original films than sequel-or-later films (6.2 out of 10). This same result is also seen in the graph, where the median of the IMDb ratings for original films is slightly higher than that of later films. It's also important to note that the distribution of ratings for sequel-or-later films seems to have greater variability, which suggests less consistent audience reception; additionally, while both distributions have outliers, the sequel-or-later films have 4 more outliers than the original films distribution, and they are also lower, suggesting that sequel-or-later films tend to have a harder time maintaining high audience enjoyment compared to original films. Nevertheless, both boxplots do overlap indicating that some sequels can still be successful.

Overall, the graph and summary statistics suggest original films have slightly greater IMDb ratings than sequel-or-later films.

### Returns by Installment

Grouping all post-original films together may hide a pattern: do returns
decline with each successive installment?

```{r}
#| label: installment-returns

final_joined_data |>
  mutate(installment = case_when(
    order == 1 ~ "1st (Original)",
    order == 2 ~ "2nd",
    order >= 3 ~ "3rd+"
  )) |>
  group_by(installment) |>
  summarize(
    median_ratio = median(income_budget_ratio, na.rm = TRUE),
    median_rating = median(im_db_rating_cleaned, na.rm = TRUE),
    n = n()
  ) |>
  knitr::kable(digits = 2)
```

The results reveal a divergence between audience reception and financial
performance. Median IMDb ratings decline steadily with each installment
(6.9 → 6.4 → 6.1), consistent with franchise fatigue. Financial returns,
however, do not follow: while second installments show a meaningful drop
in income-to-budget ratio (4.01 → 3.30), third-and-later installments
recover to near-original levels (3.96) — despite being the worst-rated
group. By the third film, a franchise's earning power appears to have
decoupled from its critical reception: audiences continue paying for
films they rate progressively lower.

One caveat: the third-and-later group is subject to survivorship bias.
Franchises only reach a third installment if earlier films performed
well enough to justify one, so this group over-represents commercially
successful series. Even so, the divergence between the ratings and
returns trajectories suggests that established franchises retain
box-office pull independent of per-film quality.

### Outlier Sequels: The Overperformers

To put faces on these patterns, we examine the individual sequels with the
highest income-to-budget ratios.

```{r}
#| label: overperformer-sequels

final_joined_data |>
  filter(order >= 2) |>
  arrange(desc(income_budget_ratio)) |>
  select(title, year_sequels, order,
         income_budget_ratio, im_db_rating_cleaned) |>
  head(10) |>
  rename(
    "Title" = title,
    "Year" = year_sequels,
    "Installment" = order,
    "Income/Budget" = income_budget_ratio,
    "IMDb Rating" = im_db_rating_cleaned
  ) |>
  knitr::kable(digits = 2)
```

The overperformer list is dominated by low-budget horror. *Paranormal
Activity 2* turned a \$3 million budget into \$177 million — a 59x
return — while its audience rated it just 5.6. Its successor returned
41x, and the fourth installment, rated a scornful 4.6, still returned
28x. Horror sequels cost little to produce and their audiences show up
regardless of how they rated the previous film, which is a large part of
what drives the third-installment recovery seen above. *The Twilight
Saga: New Moon* offers perhaps the starkest single example of the
ratings-revenue divorce: rated 4.7 on IMDb, it grossed \$711 million.
And the 2018 *Halloween* — the eleventh film in a forty-year-old
franchise — returned roughly 25x its \$10 million budget, illustrating
how durable franchise earning power can be across decades of declining
ratings.

Note on data quality: two entries in this table (*Hellraiser*, 1987, and
*Gremlins*, 1984) appear to reflect join mismatches, where a sequel's row
matched the original film's financial data within the 2-year tolerance
window. These entries are retained in the table for transparency but are
excluded from interpretation.

### Outlier Sequels: The Flops

The least profitable sequels tell an equally instructive story.

```{r}
#| label: flop-sequels

final_joined_data |>
  filter(order >= 2) |>
  arrange(income_budget_ratio) |>
  select(title, year_sequels, order,
         income_budget_ratio, im_db_rating_cleaned) |>
  head(10) |>
  rename(
    "Title" = title,
    "Year" = year_sequels,
    "Installment" = order,
    "Income/Budget" = income_budget_ratio,
    "IMDb Rating" = im_db_rating_cleaned
  ) |>
  knitr::kable(digits = 2)
```

Sequels do fail — but the pattern of failure is revealing. *Basic
Instinct 2* spent \$70 million arriving fourteen years after the
original; *Zoolander 2* waited fifteen; *RoboCop 3* recast its lead.
The common thread across the flops is not simply poor quality — several
profitable sequels in the dataset were rated worse — but broken
continuity: long gaps between installments, missing stars, or franchises
whose momentum had lapsed. This suggests the decoupling of earnings from
reception is conditional: franchise loyalty survives declining quality,
but not lapsed momentum. (Note also that the very lowest ratios, such as
*Universal Soldier: Regeneration*, partly reflect releases that went
largely direct-to-video, where theatrical gross understates total
revenue.)

### Predicting Sequel Performance

The descriptive results raise a practical question: if a studio has an
original film in hand, can its characteristics predict how a sequel would
perform? To explore this, we restructure the data so that each observation
is a franchise pair — the original film's characteristics alongside its
second installment's outcome — and fit a linear regression predicting the
sequel's income-to-budget ratio.

```{r}
#| label: build-pairs
#| message: false
#| warning: false

pairs <- final_joined_data |>
  filter(order <= 2) |>
  group_by(movie_series, order) |>
  slice_min(year_diff, n = 1, with_ties = FALSE) |>  # resolve duplicate title matches
  ungroup() |>
  select(movie_series, order, year_sequels,
         income_budget_ratio, im_db_rating_cleaned,
         budget) |>
  pivot_wider(
    id_cols     = movie_series,
    names_from  = order,
    values_from = c(income_budget_ratio, im_db_rating_cleaned,
                    budget, year_sequels)
  ) |>
  rename(
    original_ratio   = income_budget_ratio_1,
    sequel_ratio     = income_budget_ratio_2,
    original_rating  = im_db_rating_cleaned_1,
    original_budget  = budget_1,
    year_original    = year_sequels_1,
    year_sequel      = year_sequels_2
  ) |>
  mutate(gap_years = year_sequel - year_original) |>
  filter(!is.na(sequel_ratio), !is.na(original_ratio),
         is.finite(sequel_ratio), is.finite(original_ratio),
         original_budget > 0)
```

Complete data for both halves of a pair survives the joins for
`r nrow(pairs)` franchises. We model the log of the sequel's
income-to-budget ratio as a function of four characteristics of the
original: its own income-to-budget ratio (logged), its IMDb rating, its
production budget (logged), and the number of years between the original
and the sequel. Monetary variables are logged because return ratios are
heavily right-skewed — a small number of extreme performers would
otherwise dominate the fit.

```{r}
#| label: sequel-model

model <- lm(log(sequel_ratio) ~ log(original_ratio) + original_rating +
              log(original_budget) + gap_years,
            data = pairs)

broom::tidy(model) |>
  mutate(term = case_when(
    term == "(Intercept)"          ~ "Intercept",
    term == "log(original_ratio)"  ~ "Original's income/budget (log)",
    term == "original_rating"      ~ "Original's IMDb rating",
    term == "log(original_budget)" ~ "Original's budget (log)",
    term == "gap_years"            ~ "Years between films"
  )) |>
  rename(
    "Predictor" = term,
    "Estimate" = estimate,
    "Std. Error" = std.error,
    "t" = statistic,
    "p-value" = p.value
  ) |>
  knitr::kable(digits = 3)
```

```{r}
#| label: sequel-model-fit

broom::glance(model) |>
  select(r.squared, adj.r.squared, nobs) |>
  rename(
    "R-squared" = r.squared,
    "Adjusted R-squared" = adj.r.squared,
    "N (franchise pairs)" = nobs
  ) |>
  knitr::kable(digits = 3)

plot(model, which = 1)
```

The model explains roughly 60% of the variation in sequel returns
(adjusted R² = 0.60), and its results reinforce the decoupling theme from
the installment analysis:

**The original's profitability is the dominant predictor.** The
coefficient on the original's (logged) income-to-budget ratio is 0.53
(p < 0.001). Because both variables are logged, this is interpretable as
an elasticity: a 1% more profitable original predicts a roughly 0.53%
more profitable sequel. Notably, the elasticity is well below 1 —
sequels inherit their parent's economics, but with substantial regression
to the mean. Exceptional originals tend to produce good, but less
exceptional, sequels.

**The original's audience rating predicts nothing.** Controlling for the
original's profitability, its IMDb rating has essentially no relationship
with sequel returns (coefficient ≈ 0.02, p = 0.89). This is the
regression-based counterpart of the installment finding: how much
audiences *liked* the original carries no measurable information about
how much its sequel will *earn*, once the original's financial
performance is known. A studio deciding whether to greenlight a sequel
would, on this evidence, learn more from the original's balance sheet
than from its reviews.

**Larger-budget originals predict better sequel returns.** The
coefficient on the original's (logged) budget is positive (0.16,
p = 0.049) — controlling for profitability, sequels of bigger-budget
originals return more per dollar. One interpretation is that budget
proxies for studio commitment: marketing infrastructure, distribution
reach, and franchise-building intent that carry over into the sequel.
However, this estimate sits at the boundary of conventional significance
in a small sample and should be treated as tentative.

**Longer gaps point in the expected direction, but are not significant.**
Each additional year between the original and its sequel is associated
with roughly 2% lower sequel returns (coefficient −0.019), consistent
with the broken-momentum pattern visible in the flop analysis (*Basic
Instinct 2*, *Zoolander 2*). With 44 pairs, however, the estimate does
not reach statistical significance (p = 0.21). We treat the momentum
hypothesis as directionally supported but unconfirmed at this sample
size.

Several caveats apply. This is an exploratory model on 44 franchise
pairs — those that both received a sequel and survived the data joins
with complete financial information — so the estimates describe
sequels that were made, not the full universe of sequels a studio might
consider. Selection is therefore built into the sample: studios
presumably declined the least promising sequels, which restricts the
range of the data. The relatively high R² also warrants caution, as both
outcome and predictors were derived from the same data pipeline, and a
small number of extreme franchises may exert outsized influence on the
fit.

```{r}
pairs[c(5, 40), ] |> select(movie_series, original_ratio, sequel_ratio, gap_years)

# and the high-fitted-value point:
pairs |> slice_max(original_ratio, n = 1) |>
  select(movie_series, original_ratio, sequel_ratio)
```
```{r}
pa_row <- which(pairs$movie_series == "Paranormal Activity")

model_check <- lm(log(sequel_ratio) ~ log(original_ratio) + original_rating +
                    log(original_budget) + gap_years,
                  data = pairs[-c(5, 40, pa_row), ])
summary(model_check)
```

### Which standalone movies should have gotten sequels??


```{r}
#| label: greenlight-simulator

all_scored_films <- imdb_full |>
  filter(!is.na(budget), !is.na(worlwide_gross_income),
         budget > 0, worlwide_gross_income > 0,
         str_detect(country, "USA")) |>          # avoid currency mismatch trap
  mutate(income_budget_ratio = worlwide_gross_income / budget) |>
  filter(is.finite(income_budget_ratio), income_budget_ratio > 0,
         budget >= 1e4)                          # drop junk micro-budget rows

# Rename to what the model expects, assume typical 3-year turnaround
candidates <- all_scored_films |>
  mutate(
    original_ratio  = income_budget_ratio,
    original_rating = weighted_average_vote,
    original_budget = budget,
    gap_years       = 3
  )

candidates |>
  mutate(predicted_sequel_ratio = exp(predict(model, newdata = candidates))) |>
  arrange(desc(predicted_sequel_ratio)) |>
  select(original_title, year, budget, income_budget_ratio,
         weighted_average_vote, predicted_sequel_ratio) |>
  head(30) |>
  rename(
    "Title" = original_title,
    "Year" = year,
    "Budget" = budget,
    "Actual Income/Budget" = income_budget_ratio,
    "IMDb Rating" = weighted_average_vote,
    "Predicted Sequel Ratio" = predicted_sequel_ratio
  ) |>
  knitr::kable(digits = 2)
```


## Discussion

  This study investigated whether movie sequels/succeeding films or original films seem to have greater success. Success was broken down into two categories — financial success and rating success. At the aggregate level, original films have greater success in both categories; at the installment level, however, the relationship between the two kinds of success comes apart in an interesting way.
  
  Financial success was calculated as an income-to-budget ratio, and the median for original films, around 4, was slightly higher than that of sequel-or-later films, about 3.7. In other words, original films, on average, earn more money relative to money invested. This suggests that even though sequel-or-later films often have a "fan base," or a group of people invested in the first film and wanting to watch more, this doesn't often lead to greater financial success. Thus, even though original films are "riskier" in the sense that filmmakers have no basis to conclude how their film might be perceived by audiences, they still tend to result in more income.
  
  IMDb ratings were used to measure rating success, and again, the median IMDb rating for original films (6.9 out of 10) was higher than that of sequel-or-later films (6.2 out of 10). Factors that can explain decreased ratings for sequel-or-later films include a decline in the novelty of the film's plot, higher expectations that the first film established leading to a sense of being let down, and the phenomenon of "franchise fatigue," where viewers no longer look forward to a continuation of a story due to issues such as decreasing quality or lack of originality.
  
  The installment-level analysis complicates the simple "originals win" story. Ratings decline monotonically across installments (6.9 → 6.4 → 6.1) — franchise fatigue is clearly visible in how audiences evaluate the films. But financial returns do not track this decline: after a dip at the second installment (3.30), third-and-later films recover to near-original levels of profitability (3.96) while being the worst-rated group in the dataset. Franchise earning power, in other words, appears to decouple from audience reception as a series matures. This provides a statistical basis for a familiar puzzle: audiences profess exhaustion with long-running franchises, yet studios continue to greenlight them — because the exhaustion registers in ratings, not in revenue.
  
  The outlier analysis suggests both the mechanism and the boundary of this decoupling. Low-budget horror franchises (*Paranormal Activity*, *Saw*) drive much of the late-installment recovery: their sequels cost little and their audiences attend reliably regardless of reception. Meanwhile, the flop list shows that sequels fail not primarily when they are bad, but when franchise momentum breaks — long gaps (*Basic Instinct 2*, *Zoolander 2*) or the loss of key elements (*RoboCop 3*). Franchise loyalty behaves less like enthusiasm and more like habit: it tolerates declining quality but not interruption.
  
  The analysis includes several limitations. First, the third-and-later installment group carries survivorship bias: franchises only reach a third film if earlier installments performed well enough to justify one, so this group over-represents commercially successful series. Some of the late-installment recovery therefore reflects selection rather than a property of sequels as such.
  
  Second, factors like genre may influence whether sequels seem to have more success than their original films, and the outlier analysis strongly suggests genre matters — horror appears to dominate the high-return sequels — but genre was not systematically controlled for. A possible improvement is to repeat the analysis stratified by genre.
  
  Third, the title-based fuzzy join, while achieving a full match rate, introduced at least two identifiable mismatches (*Hellraiser*, *Gremlins*) where sequel rows matched original films' financial data. A stricter join keyed on unique identifiers across all datasets would improve reliability.
  
  Finally, IMDb ratings may be susceptible to problems like response bias and may not be representative of the actual quality of the movie if users are voting based on other factors, such as rating movies higher if it includes their favorite actors. A possible improvement is to incorporate additional reception measures, such as critics' scores.
  
## References

[Franchise Fatigue](https://www.licenseglobal.com/movies/trending-what-is-franchise-fatigue-and-is-it-real)

"Trending: What Is Franchise Fatigue, and Is It Real?" License Global, www.licenseglobal.com/movies/trending-what-is-franchise-fatigue-and-is-it-real.

[IMDb Rating Limitations](https://www.theringer.com/2019/06/12/tv/imdb-rating-system-problems-chernobyl)

"The Problem with IMDb's Rating System." The Ringer, www.theringer.com/2019/06/12/tv/imdb-rating-system-problems-chernobyl.

[IMDb movie ratings data set](https://www.kaggle.com/datasets/sharathmudigoudr/imdb-movie-dataset-from-year-1893-to-2020/data)

Mudigoudr, Sharath. "Imdb Movie Dataset from Year 1893 to 2020." Kaggle, 30 Mar. 2024, www.kaggle.com/datasets/sharathmudigoudr/imdb-movie-dataset-from-year-1893-to-2020/data. 

[Sequel data set](https://www.kaggle.com/code/cpbdev/sequels-dataset/input)

Cpbdev. "Sequels_dataset." Kaggle, 16 Feb. 2023, www.kaggle.com/code/cpbdev/sequels-dataset/input. 
