---
title: "Sequels: More or Less Successful"
author: 
  - name: Isha Omer
  - name: Lexi Payne
  - name: Camden Chin
  - name: David Boham
format: html
---

## Introduction

This project investigates whether movie sequels are more or less successful than 
their originals. Using two datasets — a curated list of movies with sequels and 
IMDb ratings and financial data — we explore how sequel performance compares 
across metrics like ratings, box office revenue, and audience engagement. All 
data was sourced from IMDb and compiled for the STA 199 course project.

## Data Import
```{r}
#| label: load-libraries
#| message: false
#| warning: false
library(tidyverse)
library(janitor)
library(stringr)
library(lubridate)
```
```{r}
#| label: import-data
#| message: false
#| warning: false
movies_cleaned <- read_csv("data/Movies - Cleaned.csv") |> clean_names()
movies_sequels <- read_csv("data/Movies with sequels.csv") |> clean_names()
imdb_movies    <- read_csv("data/IMDb movies.csv")        |> clean_names()
imdb_ratings   <- read_csv("data/IMDb ratings.csv")       |> clean_names()
```

## Data Cleaning

After importing the raw datasets, we clean and standardize each one to ensure 
consistency before joining. This includes fixing data types, removing duplicates, 
and dropping rows with missing values in key fields. Note that `clean_names()` 
was already applied on import, so it is not repeated here.
```{r}
#| label: clean-all-datasets
#| message: false
#| warning: false

# Clean movies_cleaned
movies_cleaned <- movies_cleaned |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean movies_sequels
movies_sequels <- movies_sequels |>
  distinct() |>
  mutate(
    title = str_squish(str_to_title(title)),
    year  = as.integer(year)
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_movies
imdb_movies <- imdb_movies |>
  distinct() |>
  mutate(
    title                 = str_squish(str_to_title(title)),
    year                  = as.integer(year),
    budget                = as.numeric(str_replace_all(budget, "[^0-9.]", "")),
    usa_gross_income      = as.numeric(str_replace_all(usa_gross_income, "[^0-9.]", "")),
    worlwide_gross_income = as.numeric(str_replace_all(worlwide_gross_income, "[^0-9.]", ""))
  ) |>
  filter(!is.na(title), !is.na(year))

# Clean imdb_ratings
imdb_ratings <- imdb_ratings |>
  distinct() |>
  filter(!is.na(imdb_title_id), !is.na(weighted_average_vote))
```

## Data Joining

Here we join the two movie datasets (Data Source 1 from the project proposal) 
using a left join on both `title` and `year` to ensure we're matching the 
correct films. The suffix argument distinguishes columns that appear in both 
datasets. The result, `joined_sequel_data`, retains all rows from 
`movies_cleaned` and appends sequel information where available.
```{r}
#| label: join-sequel-data
#| message: false
joined_sequel_data <- movies_cleaned |>
  left_join(
    movies_sequels,
    by     = c("title", "year"),
    suffix = c("_cleaned", "_sequels")
  )

head(joined_sequel_data, 5)
```

The two IMDb datasets are joined using `imdb_title_id` as the key, which is a 
unique identifier shared across both files. This produces a single IMDb dataset 
`imdb_full` that contains both movie metadata and ratings information.
```{r}
#| label: join-imdb-data
#| message: false
imdb_full <- imdb_movies |>
  left_join(imdb_ratings, by = "imdb_title_id")
```

Finally, we combine both data sources into one unified dataset. Because some 
titles exist in IMDb under a different release year, we join on title only and 
then keep the closest year match within a 2-year window. This significantly 
improves the number of successful matches over a strict title + year join.
```{r}
#| label: final-join
#| message: false
#| warning: false
final_joined_data <- joined_sequel_data |>
  left_join(
    imdb_full,
    by           = "title",
    suffix       = c("_sequels", "_imdb"),
    relationship = "many-to-many"
  ) |>
  mutate(year_diff = abs(year_sequels - year_imdb)) |>
  group_by(title, year_sequels) |>
  slice_min(year_diff, n = 1) |>
  ungroup() |>
  filter(year_diff <= 2)

head(final_joined_data, 5)
```

## Join Summary

All 272 movies successfully matched with IMDb data, confirming that the 
title-based join with a 2-year tolerance worked correctly. This gives us a 
complete dataset to work with for the rest of the analysis.
```{r}
#| label: join-summary
#| message: false
final_joined_data |>
  summarise(
    matched   = sum(!is.na(imdb_title_id)),
    unmatched = sum(is.na(imdb_title_id))
  )
```
