---
title: Add project title here
subtitle: Proposal
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

-   Address ethical concerns about the data. If none are identified, state that explicitly.

## Research question

-  Do sequels generate higher box office revenue than original films after accounting for budget? 

-   A description of the research topic
Our hypothesis is that sequels will generate less revenue than original films when accounting for budget, since budgets are usually higher for sequels, but typically don't generate as much traction or popularity with audiences. We expect there to be some outliers with big film franchises such as Marvel movies or the Star Wars franchise.

-   Identify the types of variables in your research question. Categorical? Numerical?

## Glimpse of data

```{r}
#| label: load-data-1

# add code here
```

# Data 2

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

-   Address ethical concerns about the data. If none are identified, state that explicitly.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Numerical?

## Glimpse of data

```{r}
#| label: load-data-2

# add code here
```
