---
title: Add project title here
subtitle: Proposal
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.
The source of the data is kaggle

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

-   Address ethical concerns about the data. If none are identified, state that explicitly.

## Research question

-  Do sequels generate higher box office revenue than original films after accounting for budget? 

- The research topic that our team will be looking into is if sequels generate higher revenues than original films. It's a topic that is of interest to all of us, and we aren't sure what the answer is actually going to be. Since sequels often have much larger production budgets, simply comparing total revenue might not be a fair comparison. A movie with a huge budget might naturally bring in more revenue because of marketing, special effects, or brand recognition. Therefore, we'll be dividing revenue/budget to create a revenue-to-budget ratio. This will allow us to compare how successful films are relative to how much money was invested in them. Moreover, we are also interested in potentially evaluating the IMDB rating and seeing how that differs across the different types of films.

- Our hypothesis is that sequels will generate less revenue than original films when accounting for budget, since budgets are usually higher for sequels, but typically don't generate as much traction or popularity with audiences. In other words, their revenue-to-budget ratio may be lower. We expect there to be some outliers with big film franchises such as Marvel movies or the Star Wars franchise, which could have sequels that are more successful than their original films.

- There are 4 main variables that we will be evaluating: sequel status, revenue, budget and revenue/budget ratio. Sequel status is a categorical variable, since the answer is either yes or no. Revenue is a numerical continuous variable since it can have a wide range of values and is measured in dollars. Budget is also a numerical continuous variable since it can have a wide range of values. Revenue/budget ratio is also a numerical (continuous) variable for the same reasons. A potential fifth variable is IMDB rating, which is also numerical (continuous), since it can have many different values ranging from 1-10(including decimals) and not just a select few options. 

## Glimpse of data

```{r}
#| label: load-data-1

# add code here
```

# Data 2

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

-   Address ethical concerns about the data. If none are identified, state that explicitly.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Numerical?

## Glimpse of data

```{r}
#| label: load-data-2

# add code here
```
