---
title: Movie Budgets, Revenue, and Sequels
subtitle: Proposal
project:
  type: website
  output-dir: docs
  resources:
    - presentation.pdf

author:
  - name: Isha Omer
  - name: Lexi Payne
  - name: Camden Chin
  - name: David Boham

website:
  title: ggplot
  navbar:
    left:
      - text: Home
        href: index.qmd
      - text: Presentation
        href: presentation.pdf  
      - text: Proposal
        href: proposal.qmd
      - text: About
        href: about.qmd
      - text: Contract
        href: contract.qmd
    right:
      - icon: github
        href: "git@github.com:sta199-s26/" 

format:
  html:
    theme: 
      light: flatly
      dark: darkly
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

# Data 1

## Introduction and data

-   This is a Kaggle data set with IMBd movies.

-   The data was collecting by scraping from this website: https://data.world/priyankad0993. Although I'm not sure when it was created, it was last updated 3 years ago.

-   There are two csv files in this kaggle dataset. The first csv file is called "Movies-Cleaned" and contains information about the movies such as title, release date, runtime, IMDb rating, if it's part of a movie series or not, what number is it in the series, and more. The second one is called "Movies with Sequels" and focuses more on the movies that actually have sequels, providing information on their title, genre, IMDb rating, runtime, and more.

-   The only ethical concern is if the use went against the terms and services of IMDb but it is on Kaggle, so I would err on the side of not.

## Research question

-  Do sequels generate higher box office revenue than original films after accounting for budget? 

- The research topic that our team will be looking into is if sequels generate higher revenues than original films. It's a topic that is of interest to all of us, and we aren't sure what the answer is actually going to be. Since sequels often have much larger production budgets, simply comparing total revenue might not be a fair comparison. A movie with a huge budget might naturally bring in more revenue because of marketing, special effects, or brand recognition. Therefore, we'll be dividing revenue/budget to create a revenue-to-budget ratio. This will allow us to compare how successful films are relative to how much money was invested in them. Moreover, we are also interested in potentially evaluating the IMDB rating and seeing how that differs across the different types of films.

- Our hypothesis is that sequels will generate less revenue than original films when accounting for budget, since budgets are usually higher for sequels, but typically don't generate as much traction or popularity with audiences. In other words, their revenue-to-budget ratio may be lower. We expect there to be some outliers with big film franchises such as Marvel movies or the Star Wars franchise, which could have sequels that are more successful than their original films.

- There are 4 main variables that we will be evaluating: sequel status, revenue, budget and revenue/budget ratio. Sequel status is a categorical variable, since the answer is either yes or no. Revenue is a numerical continuous variable since it can have a wide range of values and is measured in dollars. Budget is also a numerical continuous variable since it can have a wide range of values. Revenue/budget ratio is also a numerical (continuous) variable for the same reasons. A potential fifth variable is IMDB rating, which is also numerical (continuous), since it can have many different values ranging from 1-10(including decimals) and not just a select few options. 

NOTE: In order to go along with this research question, we would most likely have to perform a join with another dataset, such as the the data set for our second proposal.

## Glimpse of data

```{r}
#| label: load-data-1

url_one <- "https://www.kaggle.com/api/v1/datasets/download/thedevastator/sequels-across-genres"

temp_one <- tempfile(fileext = ".zip")
download.file(url_one, temp_one, mode = "wb") # 'wb' is important for zip files on Windows/Mac

# Unzip into a 'data' folder
unzip(temp_one, exdir = "data")

movie_sequels <- read_csv("data/Movies - Cleaned.csv")
glimpse(movie_sequels)

movie_with_sequels <- read_csv("data/Movies with sequels.csv")
glimpse(movie_with_sequels)

```

# Data 2

## Introduction and data

-   This is a kaggle dataset that has all the imdb movies from 1893-2020

-   The original way it was collected is not specified in the Kaggle Dataset, yet I assume it was web-scraped with a crawler because there is data in the dataset that is not easily obtained from the premade dataset that you can download from IMDB such as the financial information like budget. I'm not exactly sure when it was created, but the dataset was last updated two years ago.

-   From looking at the two csv's that were included in the kaggle dataset, it is split into movies and ratings. The movies csv has information about the description, actors, and directors. Meanwhile, the ratings has the average rating along with votes by age group break down

-   The only ethical concern is if the crawler use went against the terms and services of IMDB but it is on Kaggle, so I would err on the side of not.

## Research question

-   Are young people contrarians? More generally, how does age correlate with popular movies. Are younger people more likely or less likely to like popular movies in comparison to older age groups?

- This research question is important to investigate, especially for producers, as youger people tend to control box office revenue, so this question is vital for predicting profitability. It also shows cultural shifts. Popular movies will be defined as movies popular during the older age groups time and the younger age groups time so time-differences can be accounted for. Popularity will be defined as ratings within each age group - how highly the younger age group rated it, and how highly the older age group rated it.

-   The topic is viewing the correlation with age and popularity of movies. I think due to the fact that before, at least from an American-centric view, there was more unity due to a lack of individualized media. Now due to more individual media such as tik tok, it may result in a less united collective identity which would means that we hypothesize younger people may be less likely to like popular movies in comparison to older age groups.

-   We will have the numerical variables of the various age groups. We also plan to measure both by gross profit which well help to show how much a movie made thus how popular, and in terms of categorical variables using the votes as another way to perceive a movies likability/popularity.

## Glimpse of data

```{r}
#| label: load-data-2

url <- "https://www.kaggle.com/api/v1/datasets/download/sharathmudigoudr/imdb-movie-dataset-from-year-1893-to-2020"

temp <- tempfile(fileext = ".zip")
download.file(url, temp, mode = "wb") # 'wb' is important for zip files on Windows/Mac

# Unzip into a 'data' folder
unzip(temp, exdir = "data")

```


```{r}
#| label: Data-2-Exploration
ratings <- read_csv("data/IMDb ratings.csv")
movies <- read_csv("data/IMDb movies.csv")

ratings |> 
  head(5)
movies |> 
  head(5)

glimpse(ratings)
glimpse(movies)

```